---
title: "FinalReport"
output: 
  html_document:
    self-contained: true
---

## Team Members

| ![Riccardo Ferrari](https://media.licdn.com/dms/image/v2/D5603AQFflEafFyd_7A/profile-displayphoto-crop_800_800/B56ZyG5l2gIAAI-/0/1771789778338?e=1773273600&v=beta&t=FdsOTImp8CNb_dXLdDQCNNXy_j8SyvSK0DxZqOiT7jI "Riccardo Ferrari") | ![Mike Ye](https://media.licdn.com/dms/image/v2/D5603AQG9mSSoNSuGYA/profile-displayphoto-shrink_800_800/profile-displayphoto-shrink_800_800/0/1718691249838?e=1773878400&v=beta&t=2luSsNSs3Qjs0-bktoZLajVcQeoDmHPqxcX2cdPo9Z4 "Mike Ye") | ![Daniel Myung](https://drive.google.com/uc?export=view&id=1oX3kGp_eOsgvzug_oAhZMJBVH7htjpfW) | ![Steven Liu](https://media.licdn.com/dms/image/v2/D5603AQENrf52ejSqrw/profile-displayphoto-shrink_800_800/B56ZNpR_MXGsAk-/0/1732638143673?e=1773273600&v=beta&t=bZ6kbfHZRhqgfTI_Boqudv1-jZAX4SIyoGw_Qg6cgnc "Steven Liu") | ![Andrea Demelas](https://drive.google.com/uc?export=view&id=1sfs8yje5_bIvdE-VDuWeBrWkE1UmgzFm) |
|---|---|---|---|---|
| **[Riccardo Ferrari](https://www.linkedin.com/in/riccardo-ferrari-2929b722a/)**<br>Financial Analyst | **[Mike Ye](https://www.linkedin.com/in/mike-ye-460923270/)**<br>Team Lead | **[Daniel Myung]()**<br>Data Scientist | **[Steven Liu](https://www.linkedin.com/in/steven-liu-554652244/)**<br>Communication Specialist | **[Andrea Demelas](https://www.linkedin.com/in/andreademelas/)**<br>Business Analyst |

# Executive Summary
414 Company's call center currently faces several challenges. Call center associates frequently face abusive customers, which leads to extreme employee turnover. Many quit within days, driving high training costs of $1000 per associate. The call center utilizes a blanket strategy in regards to outreach, with standardized talking-points from HR. 

We use data-forward methodologies to assess customer patterns, recommending a targeted approach to achieve beyond break-even profitability. Our approach utilizes supervised and unsupervised algorithmic learning models to calculate profitable target segments, comparing the results to blanket-calling (10% success). We demonstrate our work with each type of algorithm to propose the most profitable strategy.  

# Background
Telemarketing has evolved from reactive operations to data-driven hubs using analytics for efficiency and customer satisfaction. The industry has applied predictive learning techniques from computer science research to identify and recommend calls to high customer propensity profiles ([Alternyx](recommend calls to "nearest" high-propensity profiles, reducing random dialing)). In the modern age, telemarketing has shifted away from mixed-results intuition into focused, customer-focused campaigns. 

# Data Exploration

- mike data exploration
	- some metrics 
- clustering 
	- showcase clustering graphs


We began by taking a deep dive into our data. First, we observe that some metrics regarding our calls:

- 11% success rate
- -42.9% loss per call made. 

From a purely monetary perspective, with a success rate this low, the optimal course of action is to not make any calls at all until we can increase our success rate per call to at least 20%.

```
code for that line graph with red line
```
### Reading Data and Data Cleaning

```{r}
my_data = read.csv("tele.csv")

# Add new flag of whether the customer was contacted
my_data$was_contacted = ifelse(my_data$pdays == 999, 0, 1)

removes = c("X", "duration", "pdays") # duration is unknown when calling
to_be_factors = c("job", "marital", "education", "default", "housing", "loan", "contact", "month", "day_of_week", "poutcome", "y")

my_data[removes] = NULL
my_data[to_be_factors] = lapply(my_data[to_be_factors], factor)


summary(my_data)
```
We proceed to build supervised learning models with respect to predicting "y", or the propensity of customer profiles. 

We first found the data to require additional cleaning. For instance, we removed column “X” since it had no predictive meaning. We also removed "duration" to ensure a robust predictive approach, without the model being influenced by information it will not know. 

Additionally, we also factorized all other variables, such as "job", "marital", "education", into categorical variables to facilitate fair comparisons. We prepared the data-set for model training by reserving a portion of the data for training and another for verification. We verify consistency between the training and testing data by comparing their means.

### Data Visualization

**Current profitability**
```{r}
reward_per_success = 10
training_cost = 1000
base_call_cost = 1

rates = seq(0.05, 1, by = 0.05)

calls_before_quit = 1000 + pmax(0, (rates - 0.11) * 10000)
cost_per_call = base_call_cost + training_cost / calls_before_quit

percent_profit = ((rates * reward_per_success - cost_per_call) / cost_per_call) * 100

profit_df = data.frame(
  success_rate = rates,
  percent_profit = percent_profit
)

library(ggplot2)
library(scales)

ggplot(profit_df, aes(success_rate, percent_profit)) +
  geom_line(linewidth = 1) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  scale_x_continuous(labels = percent) +
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  labs(
    title = "Percent Profit per Call vs Success Rate",
    x = "Success Rate",
    y = "Percent Profit per Call"
  )
```

For our profitability calculations, we assume the following as per the specifications:
- Each call has a static cost of 1 dollar
- Employees have a 1000 dollar training cost, with on average 1000 calls before quitting, increasing by 100 calls per 1% additional call success above the current 11% and vice versa.
- The revenue from a successful call is 10 dollars


As shown in the figure above, the percent profitability per call increases approximately linearly with the success rate. Since the break even success rate is 20%, any call with more than a 20% chance of success will generate profit. Therefore, we should make any call if the customer has 20% or more probability to buy our product. 

However, this is not the case with our firm currently. This is where our profitability currently stands:
```{r}
actual_rate = mean(my_data$y == "yes")

actual_calls = 1000 + pmax(0, (actual_rate - 0.11) * 10000)
actual_cost_per_call = base_call_cost + training_cost / actual_calls

actual_profit = ((actual_rate * reward_per_success - actual_cost_per_call) / actual_cost_per_call) * 100

ggplot(profit_df, aes(success_rate, percent_profit)) +
  geom_line(linewidth = 1) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_hline(yintercept = actual_profit, color = "red", linetype = "dotted", linewidth = 1.2) +
  annotate("text",
           x = 0.8,
           y = actual_profit,
           label = paste0("Actual profit = ", round(actual_profit, 1), "%"),
           color = "red",
           vjust = 1.5) +
  scale_x_continuous(labels = percent) +
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  labs(
    title = "Percent Profit per Call vs Success Rate",
    x = "Success Rate",
    y = "Percent Profit per Call"
  )
```

Our calls have a 11% success rate, which equates to a -42.9% loss per call made. From a purely monetary perspective, with a success rate this low, the optimal course of action is to not make any calls at all until we can increase our success rate per call to at least 20%.

We believe we can achieve this by clustering our customer base into several groups, and only focus on calling those with a >20% succuss rate. We can also make a prediction model to determine if a call meets the critieria before we make the call to maximize our profit. We will explore both of these approachs and their predicted outcomes in the following sections.

For now, we seek to understand which of the factors influence a call's success rate the most in an attempt in increase our call success rates:

```{r}
library(ggplot2)
library(dplyr)
library(scales)
library(tidyr)

## We want to check whether there is a strong relationship between y and the other variables by looking at the correlation heatmap.To make this possible, we first recode y as a binary variable (0 for no, 1 for yes) and then include this binary version of y in the correlation matrix that we plot as a heatmap.

my_data$y_bin = ifelse(my_data$y == "yes", 1, 0)

num_vars = c("y_bin","age","campaign","previous", "emp.var.rate","cons.price.idx","cons.conf.idx","euribor3m","nr.employed")

num_df = my_data %>% select(all_of(num_vars))
corr = cor(num_df, use="pairwise.complete.obs")

corr_long = as.data.frame(as.table(corr)) %>%
  rename(var1 = Var1, var2 = Var2, r = Freq)

ggplot(corr_long, aes(var1, var2, fill=r)) +
  geom_tile() +
  coord_equal() + #square grid
  theme(axis.text.x = element_text(angle=45, hjust=1)) + #rotates the bottom labels to prevent overlap
  labs(title="Correlation heatmap", x=NULL, y=NULL)
```

We can clearly see from the heatmap that there is no strong correlation between a call's success rate and the other numerical variables; at least, no relationship as a whole is large enough to be easily visible in the graph. 

Despite the groups as a whole not having significant correlation with success rate, there still exists sigificant factors within each attribute that influences it. One of the most significant relationship we have found is between age and call succuess rate:

```{r}
analyze_binary_by_group <- function(data, var, target,
                                    breaks = NULL,
                                    bin_width = NULL) {
  df <- data
 
  if (!is.null(breaks)) {
    df$group_var <- cut(df[[var]], breaks = breaks, include.lowest = TRUE)
  } else if (!is.null(bin_width)) {
    rng <- range(df[[var]], na.rm = TRUE)
    df$group_var <- cut(df[[var]],
                        breaks = seq(rng[1], rng[2], by = bin_width),
                        include.lowest = TRUE)
  } else {
    df$group_var <- as.factor(df[[var]])
  }
 
  df <- df %>%
    filter(!is.na(group_var),
           !is.na(.data[[target]]))
 
  data_summary <- df %>%
    group_by(group_var) %>%
    summarise(
      n = n(),
      yes = sum(.data[[target]]),
      .groups = "drop"
    ) %>%
    rowwise() %>%
    mutate(
      p = yes/n,
      lower = ifelse(n > 1, prop.test(yes, n)$conf.int[1], NA),
      upper = ifelse(n > 1, prop.test(yes, n)$conf.int[2], NA)
    )
 
  plot <- ggplot(data_summary, aes(x = group_var, y = p)) +
    geom_col(fill = "steelblue") +
    geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
    geom_hline(yintercept = 0.2,
               color = "red",
               linetype = "dashed",
               linewidth = 1) +
    labs(title = paste("Proportion of", target, "by", var),
         x = var,
         y = "Proportion (95% CI)") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
 
  print(plot)
 
  return(data_summary) #doesn't really matter
}


```

```{r}
v="age"
analyze_binary_by_group(
    data   = my_data,
    var    = v,
    target = "y_bin",
    breaks = seq(min(my_data[[v]], na.rm = TRUE),
                 max(my_data[[v]], na.rm = TRUE),
                 length.out = 10))
```
We have calculated a 95% confidence interval for the true value of each subgroup of the age data as indicated by the black bars on the graph. A confidence interval above the red break-even line indicates that the subgroup is profitable to call. 

As we can see, every group above the age of 62 has a more than **40%** call success rate. This makes them highly valuable customers for our calls. We should priotize calling these people before all other groups. The other age bracket of interest is from 17-26, with a call success rate confidence interval including the  break-even threshold of 20%. We should also pay attention to this group as they have the highest *potential* to become profitable and be our future customers. 

```{r}
v="previous"
analyze_binary_by_group(
    data   = my_data,
    var    = v,
    target = "y_bin",
    breaks = seq(min(my_data[[v]], na.rm = TRUE),
                 max(my_data[[v]], na.rm = TRUE),
                 length.out = 8))
```

Repeated previous contact also seems to be a significant contributor to the success rate. There seem to be a high correlation between multiple previous calls and success rate. This doesn't necessarily indicate that we should repeatably call the same unenthusiastic customer but rather use more resources on calling repeat customers or those who seem to be amenable to our product. 


```{r}
good_num_vars = c("euribor3m")

for (v in good_num_vars) {
  analyze_binary_by_group(
    data   = my_data,
    var    = v,
    target = "y_bin",
    breaks = seq(min(my_data[[v]], na.rm = TRUE),
                 max(my_data[[v]], na.rm = TRUE),
                 length.out = 10)  
  )
}
                   
```

Another interesting find was the correlation between interests rates and our call success rate. It seems lower interests rates (<3%) dramatically increase our call success rate while high interest rates (>3.57) lowers them to the point of unsuitability. This would suggest that we should increase our call output during times with lower interest rates, and vice versa.

Next, we deploy clustering, an unsupervised machine learning technique, to create nuanced customer segments within the 414 Company's telemarketing dataset to achieve profitability. Our focus is on calling those segments with a >20% succuss rate. We can also make a prediction model to determine if a call meets the criteria before we make the call to maximize our profit. 


We will explore both of these approaches and their predicated outcomes in the following sections.

## Clustering

```{r}
# my_data <- read.csv("tele.csv")

my_data <- read.csv("tele.csv", stringsAsFactors = TRUE)
# Add new flag of whether the customer was contacted
#my_data$was_contacted = ifelse(my_data$pdays == 999, 0, 1)


#removes = c("X", "duration") # duration is unknown when calling
#to_be_factors = c("job", "marital", "education", "default", "housing", "loan", "contact", "month", "day_of_week", "poutcome", "y")

my_data$X <- NULL
my_data$duration <- NULL
  
  
#my_data[to_be_factors] = lapply(my_data[to_be_factors], factor)

str(my_data)
summary(my_data)
```

```{r}
outcome <- my_data$y

outcome <- ifelse(my_data$y == "yes", 1, 0)

#turning factors into numerical values
my_data <- as.data.frame(model.matrix(~.-1, data = my_data))


minmax <- function(x){
  (x - min(x)) / (max(x) - min(x))
}


## KILLS NA VALUES
int_minmax <- as.data.frame(lapply(my_data, minmax))

set.seed(12345)
c5 <- kmeans(int_minmax, 5)
c5$size
c5$centers


## CALCULATING AVERAGE CHURN RATES FOR EACH CLUSTER

my_data$cluster_assignments <- c5$cluster

tapply(outcome, my_data$cluster_assignments, mean)

# Calculate the mean yes rate per cluster
yes_averages <- aggregate(outcome ~ my_data$cluster_assignments, FUN = mean)
colnames(yes_averages) <- c("Cluster", "Yes Rate")
yes_averages

```

# Results 

## Data Cleaning
We proceed to build supervised learning models with respect to predicting "y", or the propensity of customer profiles. 

We first found the data to require additional cleaning. For instance, we removed column “X” since it had no predictive meaning. We also removed "duration" and "pdays" to ensure a robust predictive approach, without the model being influenced by information it will not know. 

Additionally, we also factorized all other variables, such as "job", "marital", "education", into categorical variables to facilitate fair comparisons. We prepared the data-set for model training by reserving a portion of the data for training and another for verification. We verify consistency between the training and testing data by comparing their means.


## Modeling
Our approach first centered on prioritizing sensitivity, an attribute of models that minimizes false positives, which decreases the amount of non-responsive customers contacted by 414 Company. Amongst the different approaches, we discovered that KNN gave us the highest profit, though not having the highest sensitivity. 

To benchmark a model's financial performance relies firstly on calculating the success rate, that is:
```

Success Rate = True Positive / (True Positive + False Positive)
```

Then, using the success rate to quantify cost per call:
```
Training cost = 1000 / (1000 + (success_rate - baseline_success_rate) * 10000)
cost_per_call = Training Cost + 1
```

And finally calculating the profit: 

```
Profit = ($10 * True Positive) - (cost_per_call * (True Positive + False Positive))
```
We will use this framework with all models going forward. 
### Best: K Nearest Neighbor (KNN) (without clustering)
knn in danielsbranch


K Nearest Neighbor (KNN) is an algorithm that classifies new customers by finding the K most similar historical customers and predicting based on their majority outcome (buy/no-buy). In short, it creates a "customer similarity map" of where our best customers are located. 

We discovered that KNN gave us the highest profit, though not having the highest sensitivity. We hypothesize that other models, while having higher sensitivities, did not produce a sufficient amount of true positives, leading to lower profit.  

Code
```{r}
tele <- read.csv("tele.csv", stringsAsFactors = TRUE)

tele$X <- NULL
tele$duration <- NULL

tele$telepcalled <-ifelse(tele$pdays == 999, 0,1)
  
#removes = c("X", "duration") # duration is unknown when calling
#to_be_factors = c("job", "marital", "education", "default", "housing", "loan", "contact", "month", "day_of_week", "poutcome", "y")

#tele[removes] = NULL
#tele[to_be_factors] = lapply(tele[to_be_factors], factor)

str(tele)
summary(tele)
```


```{r}

tele_d <- as.data.frame(model.matrix(~ . -1, data = tele))

library(janitor)

tele_d <- clean_names(tele_d)

str(tele_d)

minmax <- function(x){
  (x - min(x)) / (max(x) - min(x))
}

#normalized data
tele_n <- as.data.frame(lapply(tele_d, minmax))

summary(tele_n)
```


### Step 3: Split Data

```{r}
set.seed(12345)
ratio <- 0.7

train_rows <- sample(1:nrow(tele_n), ratio*nrow(tele_n))

tele_train <- tele_n[train_rows, ]
tele_test <- tele_n[-train_rows, ]

```

```{r}
library(class)

#all rows, all rows except last one (y value)
#telco_train[,-ncol(telco_train)]

#similarly, all rows, all rows except last one (y value)
#telco_test[,-ncol(telco_test)]

#this works because the y value is sitting in the LAST column

#cl stands for classification label 0 or 1 (the y value)

# outputs prediction for the test data. Both builds model and uses it


tele_pred <- knn(train = tele_train[,(-ncol(tele_train)-1)],
                  test = tele_test[,(-ncol(tele_test)-1)],
                  cl = tele_train[,(ncol(tele_train)-1)],
                  k = 18)

```

### Step 6: Evaluate your prediction

```{r}
library(caret)

confusionMatrix(as.factor(tele_pred), as.factor(tele_test$y), positive = "1")
```
We chose to use k=18, because it was a relatively high k value that still gave us good sensitivity. 

```{r}
## $10 * TP - 2 *(FP + TP) = profit
# 10 = contribution margin
# 2 = cost per made per call

successrate <- sum(tele_test[, ncol(tele_test) - 1]) / nrow(tele_test)
successrate

FN <- 633
TN <- 10963

TP <- 743
FP <- 18

profitnothing <- (10 * (FN + TP)) - (2 * (TN + TP + FN + FP))

newsuccessrate <- TP / (TP + FP)
newsuccessrate

diff <- newsuccessrate - successrate
diff

newrrate <- diff * 10000 
newrrate

pricepercall <- 1000 / (newrrate + 1000)
pricepercall <- pricepercall + 1
pricepercall


profit <- (10 * TP) - (pricepercall * (TP + FP))

profit
profitnothing

## 1000 + (successrate - .11) * 10000 / 1000
```


Assuming that everyone in our model is responsive, we see: 

- success rate = 97.6%
- cost per call = $1.10
- profit = $6590.13

### KNN with clusters
knnandkmeans in danielsbranch

- 13 at $ 1315.76
- 10 at 1359.13
- 9 at 1490.16
- 7 at 919$ profit
- 5 at 1985.85 profit
- 4 at $ 2232.14
- 3 at 2216.48 profit
- 2 at 2193.67
- therefore profit is inversely correlated with number of clusters

We also trained KNN using clustering as well. In this approach, we only made predictions about customers who were in the highest propensity cluster. However, this strategy did not perform as well as KNN on its own. 

Code
```{r}

tele <- read.csv("tele.csv", stringsAsFactors = TRUE)

tele$X <- NULL
tele$duration <- NULL

tele$telepcalled <-ifelse(tele$pdays == 999, 0,1)

str(tele)
summary(tele)

# Convert categorical variables to numeric indicators
# model.matrix creates 'dummy' columns for every factor level
tele_mm <- as.data.frame(model.matrix(~ . - 1, data = tele))
```

```{r}
normalize <- function(x) {
  return ((x - min(x)) / (max(x) - min(x)))
}

# Apply normalization to the dummy-coded data
tele_norm <- as.data.frame(lapply(tele_mm, normalize))

# Assuming 'yyes' is your target after dummy coding
target <- tele_mm$yyes 

# Remove the target from the features used for clustering/training
tele_features <- tele_norm[, !names(tele_norm) %in% c("yyes", "yno")]
```

```{r}

library(class)
library(caret)

set.seed(12345)
ratio <- 0.7

train_rows <- sample(1:nrow(tele_norm), ratio*nrow(tele_norm))

tele_train <- tele_norm[train_rows, !names(tele_norm) %in% c("yyes", "yno")]
tele_test  <- tele_norm[-train_rows, !names(tele_norm) %in% c("yyes", "yno")]

train_labels <- target[train_rows]
test_labels  <- target[-train_rows]

#Generate Clusters using K-means
set.seed(12345)
tele_clusters <- kmeans(tele_norm[, !names(tele_norm) %in% c("yyes", "yno")], centers = 4)

tele_train$cluster <- tele_clusters$cluster[train_rows]
tele_test$cluster  <- tele_clusters$cluster[-train_rows]

train_data <- tele_train
test_data  <- tele_test

cluster_performance <- tapply(train_labels, train_data$cluster, mean)
golden_cluster_id <- as.numeric(names(which.max(cluster_performance)))

cat("The Best Cluster is Cluster #", golden_cluster_id, 
    " with a training success rate of ", round(max(cluster_performance)*100, 2), "%\n")

is_golden_test <- test_data$cluster == golden_cluster_id

test_data_filtered   <- test_data[is_golden_test, ]
test_labels_filtered <- test_labels[is_golden_test]

# only making predictions for the target cluster
predictions_filtered <- knn(train = train_data, 
                            test  = test_data_filtered, 
                            cl    = train_labels, 
                            k     = 18)

cm <- confusionMatrix(as.factor(predictions_filtered), 
                      as.factor(test_labels_filtered), 
                      positive = "1")

TP <- cm$table[2,2] 
FP <- cm$table[2,1] 

total_yes_baseline <- sum(test_labels == 1)
total_calls_baseline <- nrow(test_data)
baseline_success_rate <- total_yes_baseline / total_calls_baseline

# Baseline Profit (Calling everyone in the test set)
baseline_profit <- (10 * total_yes_baseline) - (2 * total_calls_baseline)

# Model Success Rate
newsuccessrate <- TP / (TP + FP)

diff <- newsuccessrate - baseline_success_rate
newrrate <- diff * 10000 

pricepercall <- (1000 / (newrrate + 1000)) + 1

profit <- (10 * TP) - (pricepercall * (TP + FP))

# Results Output
cat("New Success Rate (Precision):", round(newsuccessrate * 100, 2), "%\n")
cat("Efficiency Adjusted Price Per Call: $", round(pricepercall, 2), "\n")
cat("Total Profit from Cluster-Filtered KNN: $", round(profit, 2), "\n")

```

code

Using this method, we identify that four clusters at our constant of 18 neighbors was the most profitable: 
- success rate = 60.2%
- Cost per call = $1.17
- profit = $2232.14 dollars.

Profit appears inversely correlated with the number of clusters as well. As more clusters appear, profits decrease, with an unusual increase and subsequent plateau after eight clusters. We hypothesize that larger clusters may contain a mix of high-propensity and low-propensity customers, and after a certain number of clusters, the algorithm identifies a mostly profitable segment. 
### SVM with RBF
svm kernel testing in danielsbranch


Support Vector Machine (SVM) is a supervised algorithm that constructs a maximum-margin plane that separates categories, drawn using linear (vanilla), polynomial (poly), or 3D lines (RBF). In short, SVM finds the best dividing line between customers who buy term deposits and those who don't, making future predictions are more accurate. 

We discovered that SVM using RBF gave us the highest sensitivity amongst the SVM variants. We hypothesize that other variants such as linear (vanilla) were not able to capture the nuances of the data or were not suitable for the specific dataset. 


code
```{r}
# my_data <- read.csv("tele.csv")

tele <- read.csv("tele.csv", stringsAsFactors = TRUE)

tele$X <- NULL
tele$duration <- NULL

tele$telepcalled <-ifelse(tele$pdays == 999, 0,1)
  
#my_data[to_be_factors] = lapply(my_data[to_be_factors], factor)

str(tele)
summary(tele)
```

```{r, cache = TRUE}
ratio <- 0.7
set.seed(12345)
train_rows <- sample(1:nrow(tele), ratio*nrow(tele))

tele_train <- tele[train_rows,]
tele_test <- tele[-train_rows,]

library(kernlab)

# radial basis rbfdot
#tanhdotmodel <- ksvm(y ~ ., data = tele_train, kernel = "tanhdot")
#polydotmodel <- ksvm(y ~ ., data = tele_train, kernel = "polydot")
#vanilladotmodel <- ksvm(y ~ ., data = tele_train, kernel = "vanilladot")
#aplacedotmodel <- ksvm(y ~ ., data = tele_train, kernel = "laplacedot")
#besseldotmodel <- ksvm(y ~ ., data = tele_train, kernel = "besseldot")
#anovadotmodel <- ksvm(y ~ ., data = tele_train, kernel = "anovadot")
#splinedotmodel <- ksvm(y ~ ., data = tele_train, kernel = "splinedot")
#rbfdotmodel <- ksvm(y ~ ., data = tele_train, kernel = "rbfdot")

#saveRDS(rbfdotmodel, file = "rbf_model.rds")
#saveRDS(tanhdotmodel, file = "tanhdot.rds")
#saveRDS(polydotmodel, file = "polydot.rds")
#saveRDS(vanilladotmodel, file = "vanilla.rds")
#saveRDS(laplacedotmodel, file = "laplace.rds")
#saveRDS(besseldotmodel, file = "bessel.rds")
#saveRDS(anovadotmodel, file = "anova.rds")
#saveRDS(splinedotmodel, file = "splinedot.rds")


loaded_tan <- readRDS(file = "tanhdot.rds")
loaded_poly <- readRDS(file = "polydot.rds")
loaded_vanilla <- readRDS(file = "vanilla.rds")
loaded_laplace <- readRDS(file = "laplace.rds")
loaded_bessel <- readRDS(file = "bessel.rds")
loaded_anova <- readRDS(file = "anova.rds")
loaded_spline <- readRDS(file = "splinedot.rds")
loaded_rbf <- readRDS(file = "rbf_model.rds")


tanh_pred <- predict(loaded_tan, tele_test)
poly_pred <- predict(loaded_poly, tele_test)
v_pred <- predict(loaded_vanilla, tele_test)
l_pred <- predict(loaded_laplace, tele_test)
b_pred <- predict(loaded_bessel, tele_test)
a_pred <- predict(loaded_anova, tele_test)
s_pred <- predict(loaded_spline, tele_test)
rbf_pred  <- predict(loaded_rbf, tele_test)


#table(v_pred, tele_test$y)

## NOT A BINARY PREDICTION, so don't use confusion matrix

## this time, use confusion matrix because this is a binary prediction
library(caret)
confusionMatrix(as.factor(tanh_pred), as.factor(tele_test$y), positive = "yes")
confusionMatrix(as.factor(poly_pred), as.factor(tele_test$y), positive = "yes")
confusionMatrix(as.factor(v_pred), as.factor(tele_test$y), positive = "yes")
confusionMatrix(as.factor(l_pred), as.factor(tele_test$y), positive = "yes")
confusionMatrix(as.factor(b_pred), as.factor(tele_test$y), positive = "yes")
confusionMatrix(as.factor(a_pred), as.factor(tele_test$y), positive = "yes")
confusionMatrix(as.factor(s_pred), as.factor(tele_test$y), positive = "yes")
confusionMatrix(as.factor(rbf_pred), as.factor(tele_test$y), positive = "yes")
```
       
```{r}
FN <- 1056
TN <- 10815

TP <- 320
FP <- 166 

successrate <- (FN + TP) / (FN + TN + TP + FP)
successrate

profitnothing <- (10 * (FN + TP)) - (2 * (TN + TP + FN + FP))

newsuccessrate <- TP / (TP + FP)
newsuccessrate

diff <- newsuccessrate - successrate
diff

newrrate <- diff * 10000 
newrrate

pricepercall <- 1000 / (newrrate + 1000)
pricepercall <- pricepercall + 1
pricepercall


profit <- (10 * TP) - (pricepercall * (TP + FP))

profit
profitnothing

```


Using SVM with RBF, we see:

- success rate = 65.8%
- Cost per call = $1.15
- profit = $2638.89 dollars.

### Logistic Regression
part a in stevenliu branch

Logistic regression models the probability of binary outcomes using the logistic (sigmoid) function, incorporating customer characteristics such as age, job type, and loan status to identify how likely each customer will be responsive.  

We discovered that logistic regression gave us a comparable sensitivity to KNN. Additionally, logistic regression gave us the second most profitable, only behind KNN. We hypothesize that logistic regression may be more precise, giving us fewer but highly responsive customers. This leads to high cost per call in additional to higher margin compared to KNN. 

```{r}
my_data = read.csv("tele.csv")

# Add new flag of whether the customer was contacted
my_data$was_contacted = ifelse(my_data$pdays == 999, 0, 1)

removes = c("X", "duration", "pdays") # duration is unknown when calling
to_be_factors = c("job", "marital", "education", "default", "housing", "loan", "contact", "month", "day_of_week", "poutcome", "y")

my_data[removes] = NULL
my_data[to_be_factors] = lapply(my_data[to_be_factors], factor)


summary(my_data)
```

```{r}
ratio = 0.7
set.seed(12345)
train_rows = sample(1:nrow(my_data), ratio*nrow(my_data))
training = my_data[train_rows, ]
testing = my_data[-train_rows, ]
```

```{r, cache= TRUE}
linear_model = glm(y ~ ., data = training, family = binomial)
stepmodel = step(linear_model, direction = "backward", trace = 0)

### Logistic model creation

#full_stepmodel = step(linear_model, direction = "backward")
#saveRDS(full_stepmodel, file = "full_stepmodel_without_pdays.rds")
#loaded_full_stepmodel <- readRDS("full_stepmodel_without_pdays.rds")

library(caret)
pred_class <- predict(stepmodel, testing, type = "response")
pred_factor <- ifelse(pred_class > 0.2, "yes", "no")
pred_factor <- factor(pred_factor, levels = levels(training$y))

confusionMatrix(pred_factor, testing$y, positive = "yes")

baseline <- 0.11
success_rate_glm <- 762/(1039+762)
retention_glm <- 1000+((success_rate_glm - baseline)*10000)
whole_training_cost <- 1000

training_cost_per_call <- whole_training_cost/retention_glm

print(training_cost_per_call)
revenue <- 10*762
```

```{r}
cat("Total revenue is:", "$", revenue)

total_cost <- 1801*(1+training_cost_per_call)


cat("Total Cost is :", "$", total_cost)

profit <- 10*762 - 1801*(1+training_cost_per_call)

cat("Total Profit is:", "$", profit)

```

Our fine-tuned logistic regression at a threshold of 0.18 sees: 
- success rate = 40.3% 
- cost per call = 1.25
- profit = 5439.79 

