---
title: "KNN"
author: "Daniel Myung"
date: "2026-02-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
tele <- read.csv("tele.csv", stringsAsFactors = TRUE)

tele$X <- NULL
tele$duration <- NULL

tele$telepcalled <-ifelse(tele$pdays == 999, 0,1)
  
#removes = c("X", "duration") # duration is unknown when calling
#to_be_factors = c("job", "marital", "education", "default", "housing", "loan", "contact", "month", "day_of_week", "poutcome", "y")

#tele[removes] = NULL
#tele[to_be_factors] = lapply(tele[to_be_factors], factor)

str(tele)
summary(tele)
```


```{r}

tele_d <- as.data.frame(model.matrix(~ . -1, data = tele))

library(janitor)

tele_d <- clean_names(tele_d)

str(tele_d)

minmax <- function(x){
  (x - min(x)) / (max(x) - min(x))
}

#normalized data
tele_n <- as.data.frame(lapply(tele_d, minmax))

summary(tele_n)
```


### Step 3: Split Data

```{r}
set.seed(12345)
ratio <- 0.7

train_rows <- sample(1:nrow(tele_n), ratio*nrow(tele_n))

tele_train <- tele_n[train_rows, ]
tele_test <- tele_n[-train_rows, ]

```

```{r}
library(class)

#all rows, all rows except last one (y value)
#telco_train[,-ncol(telco_train)]

#similarly, all rows, all rows except last one (y value)
#telco_test[,-ncol(telco_test)]

#this works because the y value is sitting in the LAST column

#cl stands for classification label 0 or 1 (the y value)

# outputs prediction for the test data. Both builds model and uses it


tele_pred <- knn(train = tele_train[,(-ncol(tele_train)-1)],
                  test = tele_test[,(-ncol(tele_test)-1)],
                  cl = tele_train[,(ncol(tele_train)-1)],
                  k = 18)

```

### Step 6: Evaluate your prediction

```{r}
library(caret)

confusionMatrix(as.factor(tele_pred), as.factor(tele_test$y), positive = "1")
```
We chose to use k=18, because it was a relatively high k value that still gave us good sensitivity. 

```{r}
## $10 * TP - 2 *(FP + TP) = profit
# 10 = contribution margin
# 2 = cost per made per call

successrate <- sum(tele_test[, ncol(tele_test) - 1]) / nrow(tele_test)
successrate

FN <- 633
TN <- 10963

TP <- 743
FP <- 18

profitnothing <- (10 * (FN + TP)) - (2 * (TN + TP + FN + FP))

newsuccessrate <- TP / (TP + FP)
newsuccessrate

diff <- newsuccessrate - successrate
diff

newrrate <- diff * 10000 
newrrate

pricepercall <- 1000 / (newrrate + 1000)
pricepercall <- pricepercall + 1
pricepercall


profit <- (10 * TP) - (pricepercall * (TP + FP))

profit
profitnothing

## 1000 + (successrate - .11) * 10000 / 1000
```

Under the assumption that we call everyone that our model predicts will say yes, and while looking at this specific testing set, we make a profit of $ 6590.13. If we try using SVM with the rbf kernel we make a profit of roughlt $ 2638.89 dollars. We also tried to combine KNN with K-means clusters, where we only tried to make predictions about people who were in the cluster that had the most yesses. This method had an estimated profit of around 1919.74 dollars. We believe this combined strategy did not perform as well as well as KNN on its own simply because there were not enough people to choose from, even though the correctly predicted yesses went up.



